---
title: "PSet2_R"
author: "Hady Aly"
date: "4/20/2022"
output: pdf_document
---

## Problem 5

```{r}

df <- read.csv("ak91.csv")
yrs_educ <- df$YRS_EDUC
wkly_wage <- df$WKLY_WAGE

```

**a)**

```{r}

# Defining X:
df$X <- 1*(yrs_educ==16)

# Sample Analogue Estimator:
Px <- mean(df$X)
Px

```

**b)**

```{r}

mu_college <- mean(df$X*df$WKLY_WAGE)/Px
mu_college

```

**c)**

```{r}

# Sample Analogue Estimator for Var(Y|X=1):
var_college <- (mean(df$X*(df$WKLY_WAGE)^2)/Px)-(mean(df$X*df$WKLY_WAGE)/Px)^2


se_college <- (1/sqrt(nrow(df)))*sqrt(var_college/Px)
se_college

```

**d)**

```{r}

Cn <- c(mu_college-1.96*se_college, mu_college+1.96*se_college)
Cn

```

**e)** Given that the confidence interval calculated at the 5% significance does not include 600, we do reject the null hypothesis that mu_college is equal to 600.

This is saying that we found statistically significant evidence against the hypothesis that the average weekly wage of those with a college degree (exactly 16 years of education) is 600.

**f)** We fail to reject the null hypothesis that mu_college is equal to 595 at a 5% significance level, given that value is included in the confidence interval.

This is saying that we haven't found statistically significant evidence against the average weekly wage of those with a college degree being equal to 595, thereby won't reject the hypothesis.

## Problem 6

**a)**

```{r}

my_confint <- function(mu_hat, se, alpha) {
z <- qnorm(1-(alpha/2))
confint <- c(mu_hat-(z*se),mu_hat+(z*se))
return(confint)
}

my_confint(mu_college , se_college , 0.01)

```

**b)**

```{r}

# Define a custom function that returns TRUE if mu_0 is not in confint
my_testrejects <- function(confint,mu_0) {
# Check whether mu_0 is in confint
is_in_confint <- mu_0 >= confint[1] & mu_0<=confint[2]
# If mu_0 is in confit, don’t reject. Else, reject.
is_rejected <- !is_in_confint
# Return boolean
return(is_rejected)
}

# Check whether the test rejects on 1\% significance level
confint_01 <- my_confint(mu_college , se_college , 0.01)
my_testrejects (confint_01, 600) # [1] FALSE

# Check whether the test rejects on 10\% significance level
confint_10 <- my_confint (mu_college , se_college , 0.1)
my_testrejects(confint_10, 600) # [1] TRUE

```

**c)**

```{r}

# Define a custom function for a two-sided test
my_twosidedtest <- function(mu_hat, se, alpha, mu_0) {
# Compute the confidence interval w/ significance level alpha
confint <- my_confint(mu_hat, se, alpha)
# Check whether mu_0 is in the confidence interval
is_rejected <- my_testrejects(confint, mu_0)

# Construct test message
if(is_rejected) {
  message <- paste("We reject the null hypothesis that mu equals",mu_0,"at a",alpha, "significance level")
 } else {
  message <- paste("We fail to reject the null hypothesis that mu equals",mu_0,"at a",alpha, "significance level")
}
# Print the message
print(message)
}

# Check whether the test rejects on 1\% significance level
my_twosidedtest(mu_college , se_college , 0.01 , 600) # Should not reject

# Check whether the test rejects on 10\% significance level
my_twosidedtest(mu_college , se_college , 0.10 , 600) # Should reject


```